Are environmental and geographic variables effective surrogates for genetic variation in conservation planning?

### 2016-06-29 code/rmarkdown/article.Rmd
+ changes to article + added push and pulling to makefile + increased surrogate space targets

<!-- scenario analysis -->
To determine if the addition of environmental and geographic targets could improve the effectiveness of prioritizations in a more realistic context, we generated prioritizations according to three scenarios using different combinations of targets (Figure 3). We then computed the proportion of adaptive and neutral genetic variation secured in each prioritization (Figure 4). The proportion of genetic variation secured in a prioritization was found to depend on the combination of targets used to generate the prioritization (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$Pr..Chisq.[[2]])`)$. Specifically, the amount-based prioritizations tended to secure less variation than the prioritizations generated using additional surrogate (`r round(plogis(scenario.GLHT.2.DF$estimate[[1]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.2.DF$std.error[[1]])*100,2)` SE ; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic variables (`r round(plogis(scenario.GLHT.2.DF$estimate[[2]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.2.DF$std.error[[2]])*100,2)` SE; `$Z_1 = `r round(scenario.GLHT.2.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$). No evidence was found to suggest that the prioritizations that included additional surrogate- or genetic-based targets secured a different proportion of genetic variation (`r round(plogis(scenario.GLHT.2.DF$statistic[[3]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.2.DF$std.error[[3]])*100,2)` SE; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$). We also found that the prioritizations tended to secure more adaptive than neutral genetic variation (`r round(plogis(scenario.GLHT.3.DF$statistic[[1]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.3.DF$std.error[[1]])*100,2)` SE; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.3.DF$Pr..Chisq.[[2]])`$). We also found significant differences in the average proportion of genetic variation secured by prioritizations in each of the scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.2.DF$statistic[[2]],2)`$; $P$ `r pvalString(scenario.ANOVA.2.DF$Pr..Chisq.[[2]])`). Specifically, both the multi-species (`r round(plogis(scenario.GLHT.1.DF$estimate[[1]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.1.DF$std.error[[1]])*100,2)` SE; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$) and multi-species with cost (`r round(plogis(scenario.GLHT.1.DF$estimate[[2]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.1.DF$std.error[[2]])*100,2)` SE; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[2]])`$) scenarios were found to secure more genetic variation than the single species scenario. Neither the multi-species scenarios were found to secure a different proportion of genetic variation (`r round(plogis(scenario.GLHT.1.DF$estimate[[3]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.1.DF$std.error[[3]])*100,2)` SE; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[3]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[3]])`$). These results suggest that surrogate-based targets can substantially improve the proportion of intra-specific genetic variation secured in a prioritization for most species.

@@

<!-- neutral genetic variation secured in prioritisations-->
Generally, the addition of environmental and geographic targets were needed to obtain prioritisations that secure a representative proportion of neutral genetic variation (Figure 4). With exception to the multi-species with cost planning scenario (Figure 4h), prioritisations generated using only amount targets failed to secure a representative proportion of neutral genetic variation ($>=$ `r rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %) for most species (Figures 4b, 4f, 4h). Whereas, in all scenarios, the prioritisations generated using amount and surrogate based targets secured a representative proportion of neutral genetic variation for most species. Although the the latter prioritisation did not secure a representative sample of the neutral genetic variation in all species, this result suggests that the use of environmental and geographic targets in conservation planning can result in a more effective prioritisation. Furthermore, this improvement was associated with only a minor increase in cost (XXXX)--emphasizing the utility of surrogates in conservation.

>>>>

<!-- scenario analysis -->
To determine if the addition of environmental and geographic targets could improve the effectiveness of prioritizations in a more realistic context, we generated prioritizations according to three scenarios using different combinations of targets (Figure 3). We then computed the proportion of adaptive and neutral genetic variation secured in each prioritization (Figure 4). The proportion of genetic variation secured in a prioritization was found to depend on the combination of targets used to generate the prioritization (${\chi^2}_{`r scenario.ANOVA.4.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.4.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.4.DF$Pr..Chisq.[[2]])`)$.

Specifically, the amount-based prioritizations tended to secure less variation than the prioritizations generated using additional surrogate (`r round(plogis(scenario.GLHT.2.DF$estimate[[1]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.2.DF$std.error[[1]])*100,2)` SE ; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[1]])`$) or genetic variables (`r round(plogis(scenario.GLHT.2.DF$estimate[[2]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.2.DF$std.error[[2]])*100,2)` SE; `$Z_1 = `r round(scenario.GLHT.2.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.2.DF$p.value[[2]])`$). No evidence was found to suggest that the prioritizations that included additional surrogate- or genetic-based targets secured a different proportion of genetic variation (`r round(plogis(scenario.GLHT.2.DF$statistic[[3]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.2.DF$std.error[[3]])*100,2)` SE; $Z_1 = `r round(scenario.GLHT.2.DF$statistic[[3]],2)`$; $P = `r pvalString(scenario.GLHT.2.DF$p.value[[3]])`$).

We also found that the prioritizations tended to secure more adaptive than neutral genetic variation (`r round(plogis(scenario.GLHT.3.DF$statistic[[1]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.3.DF$std.error[[1]])*100,2)` SE; ${\chi^2}_{`r scenario.ANOVA.3.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.3.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.ANOVA.3.DF$Pr..Chisq.[[2]])`$). We also found significant differences in the average proportion of genetic variation secured by prioritizations in each of the scenarios (${\chi^2}_{`r scenario.ANOVA.2.DF$Chi.Df[[2]]`} = `r round(scenario.ANOVA.2.DF$statistic[[2]],2)`$; $P$ `r pvalString(scenario.ANOVA.2.DF$Pr..Chisq.[[2]])`).

Specifically, both the multi-species (`r round(plogis(scenario.GLHT.1.DF$estimate[[1]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.1.DF$std.error[[1]])*100,2)` SE; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[1]],2)`$; $P = `r pvalString(scenario.GLHT.1.DF$p.value[[1]])`$) and multi-species with cost (`r round(plogis(scenario.GLHT.1.DF$estimate[[2]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.1.DF$std.error[[2]])*100,2)` SE; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[2]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[2]])`$) scenarios were found to secure more genetic variation than the single species scenario. Neither the multi-species scenarios were found to secure a different proportion of genetic variation (`r round(plogis(scenario.GLHT.1.DF$estimate[[3]])*100,2)` % mean difference $\pm$ `r round(plogis(scenario.GLHT.1.DF$std.error[[3]])*100,2)` SE; $Z_1 = `r round(scenario.GLHT.1.DF$statistic[[3]],2)`$; $P `r pvalString(scenario.GLHT.1.DF$p.value[[3]])`$).

These results suggest that surrogate-based targets can substantially improve the proportion of intra-specific genetic variation secured in a prioritization for most species.

@@

<!-- neutral genetic variation secured in prioritisations-->
Generally, the addition of environmental and geographic targets were needed to obtain prioritisations that secure a representative proportion of neutral genetic variation (Figure 4). With exception to the multi-species with cost planning scenario (Figure 4h), prioritisations generated using only amount targets failed to secure a representative proportion of neutral genetic variation ($>=$ `r rapr.params.LST[[MODE]]$scenario.analysis$genetic.target*100` %) for most species (Figures 4b, 4f, 4h). Whereas, in all scenarios, the prioritisations generated using amount and surrogate based targets secured a representative proportion of neutral genetic variation for most species. Although the the latter prioritisation did not secure a representative sample of the neutral genetic variation in all species, this result suggests that the use of environmental and geographic targets in conservation planning can result in a more effective prioritisation. Furthermore, this improvement was associated with only a moderate increase in cost in both scenarios (scenario without opportunity: `r round((score(multi.spp.prioritisations[[2]])/score(multi.spp.prioritisations[[1]]))*100,2)` %; scenario with opportunity cost: `r  round((score(multi.spp.prioritisations.with.cost[[2]])/score(multi.spp.prioritisations.with.cost[[1]]))*100,2)` %)--emphasizing the utility of surrogates in conservation.
